---
title: "Reproducible Research - Course Project 1"
author: "Andrew Golus"
date: "March 30, 2017"
output:
  html_document: default
  pdf_document: default
---

### Load the dataset
```{r load}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/ag827/Desktop/R")
activity <- read.csv("activity.csv")
```

### Create histogram of the total number of steps taken each day
```{r histogram1}
library(dplyr)
steps.by.day <- activity %>%
        filter(is.na(steps)==FALSE) %>%
        group_by(date) %>%
        summarize(mean.steps = sum(steps))
hist(steps.by.day$mean.steps, main = "Histogram", xlab = "Average Number of Steps per Day")
```

### Calculate mean number of steps taken each day
```{r mean1}
mean(steps.by.day$mean.steps)
```

### Calculate median number of steps taken each day
```{r median1}
quantile(steps.by.day$mean.steps, probs = 0.5)
```

### Create time series plot of the average number of steps taken
```{r time.series}
steps.by.interval <- activity %>%
        filter(is.na(steps)==FALSE) %>%
        group_by(interval) %>%
        summarize(mean.steps = mean(steps))
plot(steps.by.interval$interval, steps.by.interval$mean.steps, type = "l", xlab = "Interval", ylab = "Average Number of Steps", main = "Series Plot")
```

### Determine the 5-minute interval that, on average, contains the maximum number of steps
```{r max.interval}
subset(steps.by.interval, steps.by.interval$mean.steps == max(steps.by.interval$mean.steps))[1,1]
```

### Calculate the total number of missing values in the dataset
```{r missing}
sum(is.na(activity$steps) == TRUE)
```

### Create a new dataset with imputed missing data
My strategy is to replace the missing step values with the average number of steps for the interval.
```{r impute}
x0 <- subset(activity, is.na(activity$steps) == FALSE)
x1 <- subset(activity, is.na(activity$steps) == TRUE)
x1 <- merge(x1, steps.by.interval, all.x = TRUE)
x1 <- x1[,-2]
names(x1) <- c("interval", "date", "steps")
activity.fix <- rbind.data.frame(x0, x1)
steps.by.day <- activity.fix %>%
        filter(is.na(steps)==FALSE) %>%
        group_by(date) %>%
        summarize(mean.steps = sum(steps))
```

### Create histogram of the total number of steps taken each day after missing values are imputed
```{r histogram2}
hist(steps.by.day$mean.steps, main = "Histogram", xlab = "Average Number of Steps per Day")
```

### Calculate mean number of steps taken each day after missing values are imputed
```{r mean2}
mean(steps.by.day$mean.steps)
```

### Calculate median number of steps taken each day after missing values are imputed
```{r median2}
quantile(steps.by.day$mean.steps, probs = 0.5)
```

### Compare the average number of steps taken per 5-minute interval across weekdays and weekends
```{r weekday}
activity.weekday <- activity.fix %>%
mutate(weekday = weekdays(as.Date(date))) %>%
mutate(isweekend = ifelse(weekday == "Saturday" | weekday == "Sunday","weekend","weekday")) %>%
group_by(interval, isweekend) %>%
summarize(steps = mean(steps))
library(ggplot2)
g <- ggplot(activity.weekday, aes(interval, steps))
g + geom_line() + facet_grid(isweekend ~ .)
```
